---
title: "CSDE 502 Winter 2021, Assignment 9"
author: "larsalmq"
contact: "larsalmq@uw.edu"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(captioner)
library(tidyverse)
library(magrittr)
library(knitr)
library(readstata13)
library(kableExtra)
library(haven)


figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")
```

___Explanation___:
This assignment is intended to give you more practice delving into the Add Health data set and in manipulating additional variables. 

___Instructions___: 

1. Make sure your Rmd file has no local file system dependencies (i.e., anyone should be able to recreate the output HTML using only the Rmd source file).
1. Make a copy of this Rmd file and add answers below each question. The code that generated the answers should be included, as well as the complete source code for the document.
1. Change the YAML header above to identify yourself and include contact information.
1. For any tables or figures, include captions and cross-references and any other document automation methods as necessary.
1. Make sure your output HTML file looks appealing to the reader.
1. Upload the final Rmd to your github repository.
1. Download [`assn_id.txt`](http://staff.washington.edu/phurvitz/csde502_winter_2021/assignments/assn_id.txt) and include the URL to your Rmd file on github.com.
1. Create a zip file from your copy of `assn_id.txt` and upload the zip file to the Canvas site for Assignment 9. ___The zip file should contain only the text file. Do not include any additional files in the zip file--everything should be able to run from the file you uploaded to github.com. Please use zip format and not 7z or any other compression/archive format.___

#
__Using the full household roster (you'll need to go back the full raw data source, [21600-0001-Data.dta](http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip)), create the following variables for each respondent. Document any decisions that you make regarding missing values, definitions, etc. in your narrative as well as in the R code.  Include a frequency tabulation and a histogram of each result.__

```{r read_in_zipfile, echo=TRUE, include=FALSE, warning=FALSE, message=FALSE}
temp <- tempfile() #set temp directory
download.file("http://staff.washington.edu/phurvitz/csde502_winter_2021/data/21600-0001-Data.dta.zip",temp)

data_temp <- unzip(temp, "21600-0001-Data.dta") # unzip file
new_data <- read_dta("21600-0001-Data.dta") # rename (if desired)
View(new_data) # inspect dataset

colnames(new_data) %<>% # set column names to lowercase to avoid unforced typing errors
  str_to_lower()
```


```{r echo=TRUE, include=FALSE, warning=FALSE, message=FALSE}
## Select variables and make sure they're numeric
household <- new_data %>% 
    select(aid,
           h1hr3a, h1hr3b, h1hr3c, h1hr3d, h1hr3e, h1hr3f, h1hr3g, 
           h1hr3h, h1hr3i,h1hr3j, h1hr3k, h1hr3l, h1hr3m, h1hr3n,
           h1hr3o, h1hr3p) %>% 
   mutate_at(
        .vars = vars(-aid),
        list(
            ~ as.numeric(.)
        )
    ) 

## convert each family member to a raw count, not categorical variable
household_size <- household %>% 
  mutate(h1hr3a = ifelse(household$h1hr3a < "95", 1, NA),
         h1hr3b = ifelse(household$h1hr3b < "95", 1, NA),
         h1hr3c = ifelse(household$h1hr3c < "95", 1, NA),
         h1hr3d = ifelse(household$h1hr3d < "95", 1, NA),
         h1hr3e = ifelse(household$h1hr3e < "95", 1, NA),
         h1hr3f = ifelse(household$h1hr3f < "95", 1, NA),
         h1hr3g = ifelse(household$h1hr3g < "95", 1, NA),
         h1hr3h = ifelse(household$h1hr3h < "95", 1, NA),
         h1hr3i = ifelse(household$h1hr3i < "95", 1, NA),
         h1hr3j = ifelse(household$h1hr3j < "95", 1, NA),
         h1hr3k = ifelse(household$h1hr3k < "95", 1, NA),
         h1hr3l = ifelse(household$h1hr3l < "95", 1, NA),
         h1hr3m = ifelse(household$h1hr3m < "95", 1, NA),
         h1hr3n = ifelse(household$h1hr3n < "95", 1, NA),
         h1hr3o = ifelse(household$h1hr3o < "95", 1, NA),
         h1hr3p = ifelse(household$h1hr3p < "95", 1, NA),
         )
  
  # run the rowSums. We can ignore aid by using an inline "select"
household_size <-  household_size %>% 
    mutate(household_sum = 
               rowSums(select(., -aid),
                       na.rm = TRUE))

## Total number of household members in dataset 
sum(household_size$household_sum)

## Histogram of household size
p<- ggplot(household_size,
       aes(x = household_sum)) +
      geom_histogram(color = "black", 
                     fill = "blue",
                     binwidth = 1) +
    labs(title="Add Health Households",
         x="Number of Household Members", 
         y = "Count")+
  theme_bw()

p 

## Create frequency table
household_size %>% 
  group_by(household_sum) %>% 
  summarise(n = n()) %>% 
  mutate(`cumulative sum` = cumsum(n),
         `%` = round(n / sum(n)*100, 1),
         `cumulative %` = cumsum(`%`)) %>% 
  kable(caption = "Distribution of Family Household Size") %>% 
  kable_styling(full_width = FALSE, position = "center")
```
##
__Total number in household__
There are `sum(household_size$household_sum)` household members of students in the dataset.

##
__Number of sisters__
```{r sisters, echo=TRUE, message=FALSE, warning=FALSE}
## Select variables and make sure they're numeric
sisters <- new_data %>% 
    select(aid,
           h1hr3a, h1hr3b, h1hr3c, h1hr3d, h1hr3e, h1hr3f, h1hr3g, 
           h1hr3h, h1hr3i,h1hr3j, h1hr3k, h1hr3l, h1hr3m, h1hr3n,
           h1hr3o, h1hr3p) %>% 
   mutate_at(
        .vars = vars(-aid),
        list(
            ~ as.numeric(.)
        )
    ) 

## convert each family member to a raw count, not categorical variable
sisters<- sisters %>% 
  mutate(h1hr3a = ifelse(sisters$h1hr3a == "8", 1, NA),
         h1hr3b = ifelse(sisters$h1hr3b == "8", 1, NA),
         h1hr3c = ifelse(sisters$h1hr3c == "8", 1, NA),
         h1hr3d = ifelse(sisters$h1hr3d == "8", 1, NA),
         h1hr3e = ifelse(sisters$h1hr3e == "8", 1, NA),
         h1hr3f = ifelse(sisters$h1hr3f == "8", 1, NA),
         h1hr3g = ifelse(sisters$h1hr3g == "8", 1, NA),
         h1hr3h = ifelse(sisters$h1hr3h == "8", 1, NA),
         h1hr3i = ifelse(sisters$h1hr3i == "8", 1, NA),
         h1hr3j = ifelse(sisters$h1hr3j == "8", 1, NA),
         h1hr3k = ifelse(sisters$h1hr3k == "8", 1, NA),
         h1hr3l = ifelse(sisters$h1hr3l == "8", 1, NA),
         h1hr3m = ifelse(sisters$h1hr3m == "8", 1, NA),
         h1hr3n = ifelse(sisters$h1hr3n == "8", 1, NA),
         h1hr3o = ifelse(sisters$h1hr3o == "8", 1, NA),
         h1hr3p = ifelse(sisters$h1hr3p == "8", 1, NA),
         )
  
# Row Sums
sisters <-  sisters %>% 
    mutate(sisters_sum = 
               rowSums(select(., -aid),
                       na.rm = TRUE))

## Total number of household members in dataset 
sum(sisters$sisters_sum)

## Histogram of household size
s <- ggplot(sisters,
       aes(x = sisters_sum)) +
      geom_histogram(color = "black", 
                     fill = "blue",
                     binwidth = 1) +
    labs(title="Add Health Households",
         x="Number of Sisters in Household", 
         y = "Count")+
  theme_bw()

s

## Create frequency table
sisters %>% 
  group_by(sisters_sum) %>% 
  summarise(n = n()) %>% 
  mutate(`cumulative sum` = cumsum(n),
         `%` = round(n / sum(n)*100, 1),
         `cumulative %` = cumsum(`%`)) %>% 
  kable(caption = "Distribution of Sisters in Household") %>% 
  kable_styling(full_width = FALSE, position = "center")
```
Students in the Add Health dataset report having `sum(sisters$sisters_sum)` sisters living in their households.

##
__Number of brothers__
```{r brothers, echo=TRUE, message=FALSE, warning=FALSE}
## Select variables and make sure they're numeric
brothers <- new_data %>% 
    select(aid,
           h1hr3a, h1hr3b, h1hr3c, h1hr3d, h1hr3e, h1hr3f, h1hr3g, 
           h1hr3h, h1hr3i,h1hr3j, h1hr3k, h1hr3l, h1hr3m, h1hr3n,
           h1hr3o, h1hr3p) %>% 
   mutate_at(
        .vars = vars(-aid),
        list(
            ~ as.numeric(.)
        )
    ) 

## convert each family member to a raw count, not categorical variable
brothers <- brothers %>% 
  mutate(h1hr3a = ifelse(brothers$h1hr3a == "5", 1, NA),
         h1hr3b = ifelse(brothers$h1hr3b == "5", 1, NA),
         h1hr3c = ifelse(brothers$h1hr3c == "5", 1, NA),
         h1hr3d = ifelse(brothers$h1hr3d == "5", 1, NA),
         h1hr3e = ifelse(brothers$h1hr3e == "5", 1, NA),
         h1hr3f = ifelse(brothers$h1hr3f == "5", 1, NA),
         h1hr3g = ifelse(brothers$h1hr3g == "5", 1, NA),
         h1hr3h = ifelse(brothers$h1hr3h == "5", 1, NA),
         h1hr3i = ifelse(brothers$h1hr3i == "5", 1, NA),
         h1hr3j = ifelse(brothers$h1hr3j == "5", 1, NA),
         h1hr3k = ifelse(brothers$h1hr3k == "5", 1, NA),
         h1hr3l = ifelse(brothers$h1hr3l == "5", 1, NA),
         h1hr3m = ifelse(brothers$h1hr3m == "5", 1, NA),
         h1hr3n = ifelse(brothers$h1hr3n == "5", 1, NA),
         h1hr3o = ifelse(brothers$h1hr3o == "5", 1, NA),
         h1hr3p = ifelse(brothers$h1hr3p == "5", 1, NA),
         )
  
# Row Sums
brothers <-  brothers %>% 
    mutate(brothers_sum = 
               rowSums(select(., -aid),
                       na.rm = TRUE))

## Total number of household members in dataset 
sum(brothers$brothers_sum)

## Histogram of household size
b <- ggplot(brothers,
       aes(x = brothers_sum)) +
      geom_histogram(color = "black", 
                     fill = "darkorchid",
                     binwidth = 1) +
    labs(title="Add Health Households",
         x="Number of Brothers in Household", 
         y = "Count")+
  theme_bw()

b

## Create frequency table
brothers %>% 
  group_by(brothers_sum) %>% 
  summarise(n = n()) %>% 
  mutate(`cumulative sum` = cumsum(n),
         `%` = round(n / sum(n)*100, 1),
         `cumulative %` = cumsum(`%`)) %>% 
  kable(caption = "Distribution of Brothers in Household") %>% 
  kable_styling(full_width = FALSE, position = "center")
```

Participants report `sum(brothers$brothers_sum)` brothers living in their households.

##
__Total number of siblings__
```{r siblings, echo=TRUE, message=FALSE, warning=FALSE}

## Select variables and make sure they're numeric
siblings <- new_data %>% 
    select(aid,
           h1hr3a, h1hr3b, h1hr3c, h1hr3d, h1hr3e, h1hr3f, h1hr3g, 
           h1hr3h, h1hr3i,h1hr3j, h1hr3k, h1hr3l, h1hr3m, h1hr3n,
           h1hr3o, h1hr3p) %>% 
   mutate_at(
        .vars = vars(-aid),
        list(
            ~ as.numeric(.)
        )
    ) 

View(siblings)
## can't seem to figure out 'ifelse' with an "or" statement, so
## clear the "1" value

siblings <- siblings %>% 
  mutate(h1hr3a = ifelse(siblings$h1hr3a  == "1", 0, h1hr3a),
         h1hr3b = ifelse(siblings$h1hr3b  == "1", 0, h1hr3b),
         h1hr3c = ifelse(siblings$h1hr3c  == "1", 0, h1hr3c),
         h1hr3d = ifelse(siblings$h1hr3d  == "1", 0, h1hr3d),
         h1hr3e = ifelse(siblings$h1hr3e  == "1", 0, h1hr3e),
         h1hr3f = ifelse(siblings$h1hr3f  == "1", 0, h1hr3f),
         h1hr3g = ifelse(siblings$h1hr3g  == "1", 0, h1hr3g),
         h1hr3h = ifelse(siblings$h1hr3h  == "1", 0, h1hr3h),
         h1hr3i = ifelse(siblings$h1hr3i  == "1", 0, h1hr3i),
         h1hr3j = ifelse(siblings$h1hr3j  == "1", 0, h1hr3j),
         h1hr3k = ifelse(siblings$h1hr3k  == "1", 0, h1hr3k),
         h1hr3l = ifelse(siblings$h1hr3l  == "1", 0, h1hr3l),
         h1hr3m = ifelse(siblings$h1hr3m  == "1", 0, h1hr3m),
         h1hr3n = ifelse(siblings$h1hr3n  == "1", 0, h1hr3n),
         h1hr3o = ifelse(siblings$h1hr3o  == "1", 0, h1hr3o),
         h1hr3p = ifelse(siblings$h1hr3p  == "1", 0, h1hr3p),
         )
         

siblings <- siblings %>% 
  mutate(h1hr3a = ifelse(siblings$h1hr3a  == "5", 1, h1hr3a),
         h1hr3b = ifelse(siblings$h1hr3b  == "5", 1, h1hr3b),
         h1hr3c = ifelse(siblings$h1hr3c  == "5", 1, h1hr3c),
         h1hr3d = ifelse(siblings$h1hr3d  == "5", 1, h1hr3d),
         h1hr3e = ifelse(siblings$h1hr3e  == "5", 1, h1hr3e),
         h1hr3f = ifelse(siblings$h1hr3f  == "5", 1, h1hr3f),
         h1hr3g = ifelse(siblings$h1hr3g  == "5", 1, h1hr3g),
         h1hr3h = ifelse(siblings$h1hr3h  == "5", 1, h1hr3h),
         h1hr3i = ifelse(siblings$h1hr3i  == "5", 1, h1hr3i),
         h1hr3j = ifelse(siblings$h1hr3j  == "5", 1, h1hr3j),
         h1hr3k = ifelse(siblings$h1hr3k  == "5", 1, h1hr3k),
         h1hr3l = ifelse(siblings$h1hr3l  == "5", 1, h1hr3l),
         h1hr3m = ifelse(siblings$h1hr3m  == "5", 1, h1hr3m),
         h1hr3n = ifelse(siblings$h1hr3n  == "5", 1, h1hr3n),
         h1hr3o = ifelse(siblings$h1hr3o  == "5", 1, h1hr3o),
         h1hr3p = ifelse(siblings$h1hr3p  == "5", 1, h1hr3p),
         )


siblings <- siblings %>% 
  mutate(h1hr3a = ifelse(siblings$h1hr3a  == "8", 1, h1hr3a),
         h1hr3b = ifelse(siblings$h1hr3b  == "8", 1, h1hr3b),
         h1hr3c = ifelse(siblings$h1hr3c  == "8", 1, h1hr3c),
         h1hr3d = ifelse(siblings$h1hr3d  == "8", 1, h1hr3d),
         h1hr3e = ifelse(siblings$h1hr3e  == "8", 1, h1hr3e),
         h1hr3f = ifelse(siblings$h1hr3f  == "8", 1, h1hr3f),
         h1hr3g = ifelse(siblings$h1hr3g  == "8", 1, h1hr3g),
         h1hr3h = ifelse(siblings$h1hr3h  == "8", 1, h1hr3h),
         h1hr3i = ifelse(siblings$h1hr3i  == "8", 1, h1hr3i),
         h1hr3j = ifelse(siblings$h1hr3j  == "8", 1, h1hr3j),
         h1hr3k = ifelse(siblings$h1hr3k  == "8", 1, h1hr3k),
         h1hr3l = ifelse(siblings$h1hr3l  == "8", 1, h1hr3l),
         h1hr3m = ifelse(siblings$h1hr3m  == "8", 1, h1hr3m),
         h1hr3n = ifelse(siblings$h1hr3n  == "8", 1, h1hr3n),
         h1hr3o = ifelse(siblings$h1hr3o  == "8", 1, h1hr3o),
         h1hr3p = ifelse(siblings$h1hr3p  == "8", 1, h1hr3p),
         )

siblings <- siblings %>% 
  mutate(h1hr3a = ifelse(siblings$h1hr3a  == "1", 1, 0),
         h1hr3b = ifelse(siblings$h1hr3b  == "1", 1, 0),
         h1hr3c = ifelse(siblings$h1hr3c  == "1", 1, 0),
         h1hr3d = ifelse(siblings$h1hr3d  == "1", 1, 0),
         h1hr3e = ifelse(siblings$h1hr3e  == "1", 1, 0),
         h1hr3f = ifelse(siblings$h1hr3f  == "1", 1, 0),
         h1hr3g = ifelse(siblings$h1hr3g  == "1", 1, 0),
         h1hr3h = ifelse(siblings$h1hr3h  == "1", 1, 0),
         h1hr3i = ifelse(siblings$h1hr3i  == "1", 1, 0),
         h1hr3j = ifelse(siblings$h1hr3j  == "1", 1, 0),
         h1hr3k = ifelse(siblings$h1hr3k  == "1", 1, 0),
         h1hr3l = ifelse(siblings$h1hr3l  == "1", 1, 0),
         h1hr3m = ifelse(siblings$h1hr3m  == "1", 1, 0),
         h1hr3n = ifelse(siblings$h1hr3n  == "1", 1, 0),
         h1hr3o = ifelse(siblings$h1hr3o  == "1", 1, 0),
         h1hr3p = ifelse(siblings$h1hr3p  == "1", 1, 0),
         )

# Row Sums
siblings <-  siblings %>% 
    mutate(siblings_sum = 
               rowSums(select(., -aid),
                       na.rm = TRUE))

## Total number of household members in dataset 
sum(siblings$siblings_sum)

## It would have been way faster to do this without figuring out a loop
sum(new_data$h1hr3a == 5)
sum(new_data$h1hr3b == 8) ## and continue all the way down the list

## Histogram of household size
b <- ggplot(siblings,
       aes(x = siblings_sum)) +
      geom_histogram(color = "black", 
                     fill = "darkorchid",
                     binwidth = 1) +
    labs(title="Add Health Households",
         x="Number of Siblings in Household (total)", 
         y = "Count")+
  theme_bw()

b

## Create frequency table
siblings %>% 
  group_by(siblings_sum) %>% 
  summarise(n = n()) %>% 
  mutate(`cumulative sum` = cumsum(n),
         `%` = round(n / sum(n)*100, 1),
         `cumulative %` = cumsum(`%`)) %>% 
  kable(caption = "Distribution of Households with Siblings") %>% 
  kable_styling(full_width = FALSE, position = "center")
```
Participants report having a total of `sum(siblings$siblings_sum)` siblings in the household.

#
__What proportion of students live with two biological parents? Include the analysis in your R code.__
```{r dual_parent, echo=TRUE, message=FALSE, warning=FALSE}

```


#
__Calculate the number of household members that are NOT biological mother, biological father, full brother or full sister. Create a contingency table and histogram for this variable.__

```{r non_biological, echo=TRUE, message=FALSE, warning=FALSE}

```

## Source code
```{r comment=''}
cat(readLines(con = "csde502_2021_assignment09.Rmd"), sep = '\n')
```
