---
title: "CSDE 502 - HW 10 - Mortality.org - Belgium"
author: "Lars Almquist"
date: "3/17/2021"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r warning=FALSE, message=FALSE}
library(captioner)
suppressMessages(library(tidyverse))
library(magrittr)
library(knitr)
library(kableExtra)
library(haven)
library(bookdown)

figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")
```
```{r}

```

d) Include at least one of each of the following
• text with imbedded R
• hyperlink
• table
• plot

Follow the standard scientific paper format, with these sections:
• Introduction (brief, describing the country)
• Methods
    o Data (b) Verbiage that describes the data and references the HMD data source.)
    o Analysis (c) An analysis of the data to describe something interesting about the country.)
• Results
• Discussion
• Conclusions
• References (if any)
```
# Introduction
