---
title: "CSDE 502 - Homework 4"
author: "Lars Almquist - larsalmq@uw.edu"
date: "2/3/2021"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(knitr)
library(kableExtra)
library(bookdown)
library(psych)
```


Homework #4 begins with a re-creation of a histogram of the pain of graduate school as reported by graduate students themselves.

```{r recreate_histogram, echo=FALSE, warning=FALSE, message=FALSE}
gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10),
              rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))

barplot(table(gradpain), las=1, ylab="Frequency", xlab="Pain Scale", 
        main="How painful is graduate school?")
```


```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
describe(gradpain)
## n = 94
samplesize <- 94
```
# How many graduate students are in the sample?
There are `r samplesize` graduate students in the sample.

# What is the sample mean?
The mean pain level of grad students in the sample is `r mean(gradpain)`.

```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
mean(gradpain)
```

# What is the mean of these 100 means?

```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
# Box 1
## sample size = 94
## mean = 5.5
## sd = 2.71

function_box1 <- function(x) {
  d.vec <- rnorm(n = 94, mean = 5.5, sd = 2.71) %>%
  round(2)
}

mean(function_box1(gradpain))

sd(function_box1(gradpain), round(2))
```
The mean of these 100 means is `r mean(function_box1(gradpain), round(1))`.

# What is the standard deviation of these 100 means?
The standard deviation of these 100 means is `r sd(function_box1(gradpain), round(2))`. {I do not know why the "round" command isn't working here - thoughts?}

```{r echo=FALSE, warning=FALSE, message=FALSE, results=FALSE}
## Box 2
# Use set.seed(7) then run your function passing in the "gradpain" vector calculated above and a sample size of length(gradpain). Use a loop to do this 100 times and store all 100 returned means.
set.seed(7)
function_box1(gradpain)
mean(function_box1(gradpain))
print(function_box1(gradpain))

## calculate standard deviation
sd(function_box1(gradpain), round(2))

function_mean <- mean(function_box1(gradpain))
function_mean

stdev <- sd(function_box1(gradpain), round(2))
stdev

# Neither of these are working:
## confint.default(function_box1(gradpain))
## confint(function_box1(gradpain))
```
# What does your function return for the mean of means?
According to my function, the mean of means is `r mean(function_box1(gradpain), round(2))`.

# What does your function return for the standard deviation of means?
According to my function, the standard deviation of means is `r sd(function_box1(gradpain), round(2))`. {I do not know why the "round" command isn't working here - this is the second time it has happened when computing standard deviations, even though it works fine when computing means - thoughts?}

# What does your function return for the 95% confidence interval around the mean?
Honestly not sure why the `confint` command isn't working here.
